---
title: "Plot_growth_curves_calc_growth_rates"
author: "Dai Le"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **Plot Growth Curves of different bacterial strains with GGPLOT in R**
This notebook presents functions to load and plot bacterial growth data.
log base 10 of OD600 is plotted as a function of time.
Growth rates are calculated as (ln(2)/DoublingTime)*60
For simplicity, growth rates and doubling time are calculated in a different 
process. The spreadsheet only includes final values.
```{r}
#Import packages
library(tidyverse)
library(readxl)
library(plotly)
library(openxlsx)
library(patchwork)
library(gridExtra)
library(ggprism)
library(scales)
library(emojifont)
library(Cairo)
library(ggplot2)
library(ggnewscale)
library(radiant.data)
library(stats)
library(hrbrthemes)
```
```{r}
# Check file existence and download
CheckAndDownload<-function(file1,url,destfile){
  if (file.exists(file1)){
    print('File exists, will delete')
    file.remove(file1)
    downloaded<-download.file(url,destfile,method='auto',quiet=FALSE,mode='wb')
  }else {
    downloaded<-download.file(url,destfile,method='auto',quiet=FALSE,mode='wb')
  }
}

```

```{r}
# Function to select data based on genotype
SelectData<-function(dataframe,genotype){
  dataframe[(dataframe$Genotype == genotype), ]
}
```

```{r}
# Function to plot growth curves
Plot<-function(data,x,y,color_f,genotype,k){
# Generate values to show minor ticks in log base 10
tval <- sort(as.vector(sapply(seq(1,9), function(x) x*10^seq(-3,1)))) 
ttxt <- rep("",length(tval))
ttxt[seq(1,37,9)] <- as.character(tval)[seq(1,37,9)]
# Plot function
plot<-ggplot(data,aes(x,y))+
        geom_point(aes(color=color_f,
                       fill=color_f),
                       size=4,
                       stroke=1.5)+
        scale_x_continuous(name="Time(m)",
                           expand=c(0,0),
                           limits=c(0,400))+
        scale_y_log10(name="OD600(Å)",
                      breaks=c(0.01,0.1,1),
                      limits=c(0.005,2),
                      guide="prism_minor",
                      minor_breaks=tval,
                      expand=c(0,0))+
        scale_color_prism(palette="prism_dark")+
        theme_prism(palette = "prism_dark", base_size = 16)+
        labs(title=genotype)+
        theme(legend.position = c(.1,.9),
              legend.key.height = unit(10, "pt"),
              legend.title = element_blank(),
              text=element_text(size=22))+
        # Include average growth rate values in the plot
        geom_text(x=350, y=0.01, label=paste("k~",k), size=7)
plot
}

```

```{r}
# Main function to integrate all functions
main<-function(file1,url,destfile){
  # First download file
  CheckAndDownload(file1,url,destfile)
  # From file, create 2 dataframes for growth curves and growth rates data
  df<-read_xlsx(destfile,sheet="GrowthCurves", col_names = TRUE)
  df2<-read_xlsx(destfile,sheet="GrowthRates", col_names = TRUE)
  # Calculate average growth rates
  dfd<-df2 %>% group_by(Strain,Genotype) %>%
                    summarise(avg_k=format(round(mean(GrowthRate), 2), nsmall = 2))
  # Create a list of individual plots
  p<-list()
  # Iterate through all genotypes
  for (i in unique(df$Genotype)){
    png(filename = paste(i,"GrowthCurve.png"), width = 1280, height = 800,
      units = "px", pointsize = 20, bg = "white", res =120,
      restoreConsole = TRUE)
    # Create index to plot curves 1 by 1 
    for (j in 1:length(unique(df$Genotype))){
      # Select dataframes for growth curves and growth rates data
      dff<-SelectData(dataframe=df,genotype=i)
      dfdf<-SelectData(dataframe=dfd,genotype=i)
      # Plot data
      p[[j]]<-Plot(data=dff,
                   x=dff$Time_m,
                   y=dff$OD600,
                   color_f=factor(dff$Date),
                   genotype=i,
                   k=dfdf$avg_k[dfdf$Genotype == i])
     
    }
    # Print plots and turn off dev
    print(p[[j]])
    dev.off()
  }
}

```

```{r}
#Instantiate main function with information on url and file names
url<-"https://www.dropbox.com/s/8bjclc5gev6l0c8/GrowthCurves.xlsx?dl=1"
file1<-"GrowthCurves.xlsx"
destfile<-"GrowthCurves_dl.xlsx"
main(file1,url,destfile)
```







